---
title: "GENCO-2-Join via NDC and RXCUI"
author: "Nicholas Cardamone"
date: "10/3/2024"
output: html_document
---

# Goal:
* Use RxNorm concept unique identifiers (RXCUI) and NDC identifiers to map each drug in our sample to Part D formulary files.

# What this code does:
1. Loads the molecule data and extracts lists of NDA and ANDAs from each.
2. After the interim step of going to RxMix to extract the RXCUI and NDCs associated with each NDA and ANDA in our list, we load the lists, and create clean crosswalk files of NDA and ANDA to RXCUI and NDC.
3. Attempt a join of NDA or ANDA to NDC and RXCUI. Bind together all four lists. Take unique rows. This is our list of relevant Part D data, filtered by the NDC and RXCUI that are associated with the NDA and ANDAs of the target ingredients.
4. Output is relevant Medicare Part D data 2006-2023.

# Note: Joining is chunked to reduce memory burden on computer. 


```{r, Load necessary packages}
library(tidyverse)
library(dplyr)
library(pbapply)  # For progress bar
library(haven)
```

```{r, Load Molecule Data n=698}
long_mol <- read_dta("C://Users//Nick//Desktop//VA//GENCO//extra file//molecules_10.1.24_long.dta") # Downloaded from Drugs@FDA. Contact Ravi Gupta for data. 

# String pad application number and add the appropriate prefix depending if it is a branded drug (NDA) or generic drug (ANDA).
NDA_mol = long_mol %>% filter(appl_type == 1) %>% mutate(appl_no = paste0("NDA", str_pad(appl_no, 6, pad = "0")))
ANDA_mol = long_mol %>% filter(appl_type == 2) %>% mutate(appl_no = paste0("ANDA", str_pad(appl_no, 6, pad = "0")))

long_mol <- bind_rows(NDA_mol, ANDA_mol) # Overwrite long_mol with the new appl_no variable because it matches format of appl_no we'll see later.

# Create NDA and ANDA lsits.
write.csv(NDA_mol, "ndalist10-01-2024.csv")
write.csv(ANDA_mol, "andalist10-01-2024.csv")
write.csv(long_mol, "long_mol10-01-2024.csv")

```

Interim step:
1. Go to https://mor.nlm.nih.gov/RxMix/
2. Load NDA lists on File (one id per line)
3. Select Function > findRxcuiById > Add > Scope of search: Current concepts & Type of identifier: NDA.
4. Select Add another function > getAllHistoricalNDCs > Depth: NDCs ever (in)directly associated.
5. Run & download
6. Repeat steps 1-5 for ANDA list (change type of identifier to ANDA).


```{r, Load lists of RXCUI and NDCs associated with NDA, ANDA}
rxmixNDA <- read.delim("C://Users//Nick//Downloads//table(14).txt", header = T, sep = "\t", dec = ".")
rxmixANDA<- read.delim("C://Users//Nick//Downloads//table(15).txt", header = T, sep = "\t", dec = ".")
```

```{r, Read in full Medicare Part D dataset}
combined_data <- read.csv("C://Users//Nick//Documents//part1-full_data.csv")
```

```{r, String Pad NDCs}
combined_data$ndc <- str_pad(combined_data$ndc, 11, pad = "0")
```

```{r,  Create lists of RXCUI associated with NDA}
NDA_mol <- NDA_mol %>% select(appl_no, ingredient)
list_nda <- right_join(NDA_mol, rxmixNDA, c("appl_no"= "id")) # All this does is make sure list of ANDAs is in the list of NDAs we have at the beginning.
list_nda_ndc <- data.frame(app_no = list_nda$appl_no, ndc  = list_nda$F2_ndc) # pull out NDCs only
list_nda_ndc$ndc <- str_pad(list_nda_ndc$ndc, 11, pad = "0") # string pad ndcs
list_nda_rxcui <- data.frame(app_no = list_nda$appl_no, rxcui  = list_nda$F2_rxcui) %>% distinct() # pull out list of RXCUIs only

ANDA_mol <- ANDA_mol %>% select(appl_no, ingredient)
list_anda <- right_join(ANDA_mol, rxmixANDA, c("appl_no"= "id")) # All this does is make sure list of ANDAs is in the list of ANDAs we have at the beginning.
list_anda_ndc <- data.frame(app_no = list_anda$appl_no, ndc  = list_anda$F2_ndc) # pull out NDCs only
list_anda_ndc$ndc <- str_pad(list_anda_ndc$ndc, 11, pad = "0") # string pad ndcs
list_anda_rxcui <- data.frame(app_no = list_anda$appl_no, rxcui  = list_anda$F2_rxcui) %>% distinct() # pull out list of RXCUIs only

```

```{r, Extract distinct list of NDA or ANDA to RXCUI pairings from full data}
#subtest_ndc <- full_data #%>% group_by(formulary_id) %>% slice_sample(n=10)
subset_rxcui <- combined_data %>% select(-ndc) %>% distinct() %>% drop_na(rxcui)
```

```{r, Join NDA and ANDA to Part D data via NDC and RXCUI}
# Define chunk size
chunk_size <- 100000  # Adjust the chunk size as needed

# Function to process chunks with a merge operation
process_chunk <- function(chunk, list_data, by_col) {
  result <- merge(chunk, list_data, by = by_col, all.x = TRUE)
  result <- result %>% drop_na(app_no) #%>% select(date, formulary_id, app_no)
  return(result)
}

# Function to apply chunk processing with progress bar
apply_chunks <- function(full_data, list_data, by_col, chunk_size) {
  n <- nrow(full_data)
  indices <- seq(1, n, by = chunk_size)
  
  results_list <- pblapply(indices, function(i) {
    chunk <- full_data[i:min(i + chunk_size - 1, n), ]
    process_chunk(chunk, list_data, by_col)
  })
  
  results <- bind_rows(results_list)
  return(results)
}

# Process each merge operation in chunks with a progress bar
result1 <- apply_chunks(combined_data, list_nda_ndc, "ndc", chunk_size)
result2 <- apply_chunks(subset_rxcui, list_nda_rxcui, "rxcui", chunk_size)
result3 <- apply_chunks(combined_data, list_anda_ndc, "ndc", chunk_size)
result4 <- apply_chunks(subset_rxcui, list_anda_rxcui, "rxcui", chunk_size)

# Combine the results and remove duplicates
results <- bind_rows(result1, result2, result3, result4) %>% distinct()
```


```{r, Write results to csv}
# The final results data frame is now ready - this is a large file.
write.csv(results, "part2-data-10-01-2024.csv")

```




